Lemma 1

### 2021-03-11 notes/jessie-russian-lemma.tex
Another approach to showing Q_z measurable

\begin{document}
\maketitle
	\jessie{Meeting notes Mar 1, 2021: - Careful about pulling out expectation in bullet 2; when is inner product of RV = expectation of inner product? - Point out contradicting the second line would yield contradiction of subgradient inequality - Go slowly through the definition of measurable}
\paragraph{Setting} 
$\effdom(L) = \reals^d$, and $E_q L(u,Y) < \infty$ for all $u \in \reals^d$ and $p \in \P$ \jessie{$y \in \Y$?}

@@

\begin{definition}[Measurable (set-valued function)]
	Let $\B(\reals^d)$ is the Borel $\sigma$-algebra of $\reals^d$.
	Suppose $(\Omega, \B(\Omega))$ is a measurable space and $f:\Omega \toto \reals^d$ a multifunction taking values in the set of nonempty closed subsets of $\reals^d$.
	We say $f$ is weakly measurable if, for every open $U \subseteq \reals^d$, we have
	\begin{equation*}
	\{\omega : f(\omega) \cap U \neq \emptyset\} \in \B(\Omega)~.~
	\end{equation*}
\end{definition}

@@

Fix $z$.
Consider the set-valued function $Q_z : \Omega \to 2^{\reals^d}$ defined by $Q_z(\omega) := \{V : \inprod{z}{V(\omega)} = L'(u,\omega;z), V \in \V_u \}$. \jessie{type check: okay when you think about $V$ as a RV instead of function.}
Now, we want to take a measurable function $V \in \V_u$ that is a selection of $Q_z$ satisfying the following conditions:
\begin{enumerate}
	\item $V \in \V_u$.
	\item $\E_p V \in \E_p \partial L(u,Y)$.
	\item $\inprod{z}{\E_p V} = \E_p L'(u,Y,z)$
\end{enumerate} 
If these are true, then we can take $v := \E_p V$ and observe the result.

@@

Thus, we show $Q_z$ is measurable and closed-valued.
First, we have $Q_z$ closed-valued since $L$ is convex and therefore its directional derivative is continuous. \jessie{Being lazy; will fill in later.}

As measurable functions are closed under (countable) intersections, we can simply show $Q_z$ intersection of measurable functions; namely, since we have $Q_z(\omega) = \V_u \cap f: \omega \mapsto \partial L(u,\omega) \cap g : \omega \mapsto \{c \in \reals^d : \inprod{z}{c} = L'(u,\omega;z)\}$.

$\V_u$ is measurable by construction as it is a collection of measurable sets.
Moreover, the function $f$ is measurable by \jessie{Rockafellar: Measurable Dependence, Corollary 4.6}.
Now, we aim to show the function $g$ is measurable, which follows from measurability of $L'$ and the fact $\inprod{z}{\cdot}$ is continuous in its second argument, hence Borel, the set is measurable. \jessiet{Copying Russian argument}

As $Q_z$ is the intersection of these three functions, we conclude $Q_z$ is measurable and closed-valued, and therefore has a measurable selection $V \in \V_u$ such that $\inprod{z}{V(\omega) = L'(u,\omega;z)}$ for all $\omega \in \Omega$.

@@

3.
We want to show $\inprod{z}{v} = \E_p L'(u,Y,z)$.
This follows from linearity of expectation and construction of $V$; to see this, consider
\begin{align*}
\inprod{z}{v} &= \inprod{z}{\E_p V} \\
 &= \E_p \inprod{z}{V} \jessie{Re-evaluate}\\
 &= \int \inprod{z}{V(\omega)} dp\omega \\
 &= \int L'(u,\omega; z) dp\omega \\
 &= \E_p L'(u,Y;z)
\end{align*}

>>>>

\begin{document}
\maketitle
	\jessie{Meeting notes Mar 1, 2021: - Careful about pulling out expectation in bullet 2; when is inner product of RV = expectation of inner product? - Point out contradicting the second line would yield contradiction of subgradient inequality - Go slowly through the definition of measurable}
\paragraph{Setting} 
$\effdom(L) = \reals^d$, and $\E_p L(u,Y) < \infty$ for all $u \in \reals^d$ and $p \in \P$ \jessie{$y \in \Y$?}

@@

\begin{definition}[Measurable (multifunction)]
	%Let $\B(\reals^d)$ be the Borel $\sigma$-algebra of $\reals^d$.
	Suppose $(\Omega, B)$ is a measurable space and $f:\Omega \toto \reals^d$ a multifunction taking values in the set of nonempty closed subsets of $\reals^d$.
	We say $f$ is weakly measurable if, for every open $U \subseteq \reals^d$, we have
	\begin{equation*}
	\{\omega : f(\omega) \cap U \neq \emptyset\} \in B~.~
	\end{equation*}
\end{definition}

@@

Fix $z$.
Consider the set-valued function $Q_z : \Omega \to 2^{\reals^d}$ defined by $Q_z(\omega) := \{x \in \partial L(u,\omega) : \inprod{z}{x} = L'(u,\omega;z)\}$.
Now, we want to take a measurable function $V \in \V_u$ that is a selection of $Q_z$ satisfying the following conditions:
\begin{enumerate}
	\item $V \in \V_u$.
	\item $\E_p V \in \E_p \partial L(u,Y)$.
	\item $\inprod{z}{\E_p V} = \E_p L'(u,Y,z)$
\end{enumerate} 
If these are true, then we can take $v := \E_p V$ and observe the result.

@@

Thus, we show $Q_z$ is measurable and closed-valued.
We first have $Q_z$ is closed-valued since the directional derivative is the argmax of a linear function over a compact set, which is also closed.

In order to show that $Q_z$ is $B$-measurable, it is equivalent to show $gr(Q_z)$ is $B \otimes \B(\reals^d)$-measurable. \jessie{Kinda hand-wavy} \jessie{ASSUMPTION: $B$ is a complete $\sigma$-algebra with respect to $p$.}
Now, consider $gr(Q_z) = gr(Q) \cap gr(R)$, where $Q: \omega \mapsto \partial L(u,\omega)$ and $R : \omega \mapsto \{x \mid \inprod{z}{x} = L'(u,\omega;z)\}$.
As $Q$ is measurable \jessie{Rockafellar: Measurable Dependence, Cor 4.6}, so is $gr(Q)$.
Moreover, $gr(R)$ is measurable as it is the graph of a linear multifunction, hence measurable.
Now we can take $gr(Q_z) = gr(Q) \cap gr(R)$, which is measurable by closure under intersections, and have the result.

\bigskip
\hrulefill

\jessiet{If $R$ was compact}
As measurable functions are closed under (countable) intersections, we can simply show $Q_z$ intersection of measurable functions; namely, since we have $Q_z(\omega) = Q: \omega \mapsto \partial L(u,\omega) \cap R : \omega \mapsto \{x \in \reals^d : \inprod{z}{x} = L'(u,\omega;z)\}$.
\jessiet{Look at strong measurability or see if there's a result of closure under intersections}
\jessie{If $R$ is compact, then we are done by Castaing Valadier Prop iii.4}

Moreover, the function $Q$ is measurable by \jessie{Rockafellar: Measurable Dependence, Corollary 4.6}.
Moreover, $R$ is measurable as $L'$ is measurable and $\inprod{z}{\cdot}$ is continuous in its second argument, hence Borel.

As $Q_z$ is the intersection of these two functions, we conclude $Q_z$ is measurable by \jessie{Castaing Valadier Prop iii.4} and closed-valued, and therefore has a measurable selection $V \in \V_u$ such that $\inprod{z}{V(\omega)} = L'(u,\omega;z)$ for all $\omega \in \Omega$.

\hrulefill

@@

3.
We want to show $\inprod{z}{v} = \E_p L'(u,Y,z)$.
This follows from linearity of expectation and construction of $V$; to see this, consider
\begin{align*}
\inprod{z}{v} &= \inprod{z}{\E_p V} \\
 &= \E_p \inprod{z}{V} \jessie{\text{Use finiteness of expectation of } V}
 \end{align*}
 The expectation can be pulled out of the inner product because $V$ is finite for all $\omega \in \Omega$ as it is contained in $\partial L(u,\omega)$, which is compact as $L$ is finite and convex for all $u \in \reals^d$.
 \begin{align*}
 \E_p \inprod{z}{V} &= \int \inprod{z}{V(\omega)} dp\omega \\
 &= \int L'(u,\omega; z) dp\omega \\
 &= \E_p L'(u,Y;z)
\end{align*}
